---
title: "Monthly Projections"
date: "`r format(Sys.time(), '%d %B, %Y')`"
editor_options:
  chunk_output_type: console
output: html_document
params:
  debug:
    input: checkbox
    label: Debug Mode
    value: true
  prng_seed:
    input: numeric
    label: Random Number Seed
    value: 2711
  date_of_valuation:
    input: date
    label: Date of Valuation
    value: "2024-09-30"
  reporting_path:
    input: text
    label: Working directory for the monthly valuation process
    value: ""
  database_name:
    input: text
    label: Database file containing monthly data
  provider_server_name:
    input: text
    label: "Provider Server Name"
  provider_port:
    input: numeric
    label: "Provider Port"
    value: 1433
  provider_database_name:
    input: text
    label: "Provider Database Name"
  provider_username:
    input: text
    label: "Provider Username"
  provider_password:
    input: text
    label: "Provider Password"
    value: ""
  development_factors:
    input: text
    label: "Development Factors CSV file"
    value: ""
  exceptions_extract_dates:
    input: text
    label: "Specific Dates to Use in Source Data"
    value: ""
  exceptions_bad_reporting_months:
    input: text
    label: "Specific Experience Months to Exclude from Source Data"
    value: ""
  overrides_table:
    input: text
    label: "Specific Values to Override in Source Data"
    value: ""
  key_map:
    input: text
    label: "Key Translations CSV file"
    value: ""
  source_data_table_name:
    input: text
    label: "Name of the source data table"
    value: ""
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = params$debug)
knitr::opts_knit$set(progress = TRUE, verbose = TRUE)
print(params$debug)
```

## Monthly Projections

### Initialize Libraries and Environment

```{r, Initialize Environment}
# Check if libraries are installed and load them
pacman::p_load(
  dplyr,
  lubridate
)

proj.env <- new.env()
# set.seed(params$prng_seed) # seed the random number generator
# prevent scientific notation for cross-platform compatibility
options(scipen = 999)

proj.env$time_begin <- Sys.time()
```

### Assumptions

Treaty periods are duration of attached liability from effective date through expiration date (e.g. in WrittenPattern function).

(However, policy length values are defined per treaty in "TREATIES\$Policy.Length".)

Every policy is effective on the first of the month.

LOD Treaties

- Premium dealt with as Written Premium the same as an RA with an addition inherited Unearned Premium Reserve (UEPR) at the effective date.

-   Code does not cede back UEPR at the end of the treaty (there are conditions to how this is handled not currently coded / handled well), this is currently handled in the valuation spreadsheet.

### Functions

These are more general functions that can be applied throughout the document. In later sections, context specific functions are defined at the beginning of their respective code blocks.

```{r, Localized Functions}
# Return the numeric value of second argument if available.
# Otherwise, return the value of first argument.
# (Used to revise projected values with the reported amounts.)
datareplace <- function(fallback, new_data) {
  ifelse(is.na(new_data), as.numeric(fallback), as.numeric(new_data))
}

# Return the number of months between two dates
months_between <- function(EFF, CAL) {
  num_months <- lubridate::period(as.integer((lubridate::year(CAL) - lubridate::year(EFF)) * 12 + (lubridate::month(CAL) - lubridate::month(EFF)) + ifelse(lubridate::day(CAL + 1) == 1, 1, 0)), units = "months")
  return(num_months)
}

convert_numeric_to_period <- function(x, p) {
  i <- as.integer(x)
  y <- lubridate::period(0)
  switch(p,
    "years" = {
      y <- y + lubridate::years(i)
      y <- y + convert_numeric_to_period((x - i) * 12, "months")
    },
    "months" = {
      y <- y + lubridate::period(i, units = "months")
      y <- y + convert_numeric_to_period((x - i) * 30.44, "days")
    },
    "days" = {
      y <- y + lubridate::days(i)
      y <- y + convert_numeric_to_period((x - i) * 24, "hours")
    },
    "hours" = {
      y <- y + lubridate::hours(i)
      y <- y + convert_numeric_to_period((x - i) * 60, "minutes")
    },
    "minutes" = {
      y <- y + lubridate::minutes(i)
      y <- y + convert_numeric_to_period((x - i) * 60, "seconds")
    },
    "seconds" = {
      y <- y + lubridate::seconds(i)
      # no need to go further, we don't care about milliseconds
    }
  )
  return(y)
}

# Convert Excel serial date to date format
convert_excel_date <- function(excel_date) {
  as.Date(excel_date, origin = "1899-12-30")
}

# Function to preserve NAs while summing
sum_with_na <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else {
    return(sum(x, na.rm = TRUE))
  }
}
```

### Load Data

TREATIES contains many details regarding each treaty (e.g. Effective Date, Prior Loss Ratio, Target Participation, etc...).

DEVS are the development factors used for each treaty.

overrides_table contains confirmed revisions to the monthly submission data recorded by our data provider.

```{r, Load Data}
# Import data and ensure strings aren't recognized as factors

# Establish a connection to the Microsoft SQL Database
proj.env$sqlconn <- DBI::dbConnect(odbc::odbc(), 
          Driver = "ODBC Driver 17 for SQL Server", 
          Server = paste0(params$provider_server_name, ",", params$provider_port),
          Database = params$provider_database_name, 
          UID = params$provider_username, 
          PWD = params$provider_password
)

# Query the data from the TreatyTable
proj.env$treaty_positions <- DBI::dbGetQuery(proj.env$sqlconn, paste(readLines("TreatyTable.sql"), collapse = "\n")) %>%
  filter(EffectiveDate <= params$date_of_valuation) %>%
  rename(
    KEY = ConcatenatedKey,
  ) %>%
  rowwise() %>%
  mutate(
    TreatyLength = months_between(EffectiveDate, ExpirationDate),
    AssumedPolicyLengthMonths = convert_numeric_to_period(AssumedPolicyLengthMonths, "months"),
  ) %>%
  ungroup()

# Close the database connection
DBI::dbDisconnect(proj.env$sqlconn)

proj.env$development_factors <- read.csv(
  paste(params$reporting_path, params$development_factors, sep = ""),
  stringsAsFactors = FALSE
)

proj.env$extract_dates <- read.csv(
  paste(params$reporting_path, params$exceptions_extract_dates, sep = ""),
  stringsAsFactors = FALSE
) %>%
  select(-key) %>%
  mutate(
    extract_to_use = convert_excel_date(extract_to_use),
    nre_feed = as.character(nre_feed)
  ) %>%
  rename(extract_to_use_reason = Reason)

proj.env$experience_to_omit <- read.csv(
  paste(params$reporting_path, params$exceptions_bad_reporting_months, sep = ""),
  stringsAsFactors = FALSE
) %>%
  select(-key) %>%
  mutate(
    month_to_exclude = convert_excel_date(month_to_exclude),
    nre_feed = as.character(nre_feed)
  ) %>%
  rename(reason_to_drop = Reason)

proj.env$overrides_table <- read.csv(
  paste(params$reporting_path, params$overrides_table, sep = ""),
  stringsAsFactors = FALSE
) %>%
  mutate(
    bordereaux_month = convert_excel_date(bordereaux_month),
    nre_feed = as.character(nre_feed)
  ) %>%
  select(-key) %>%
  rename(reason_to_override = Reason)
```

### Load latest query of data and merge with the Exception Table

```{r, Query Most Recent Experience Data}
proj.env$con <- DBI::dbConnect(
  RSQLite::SQLite(),
  dbname = paste(params$reporting_path, params$database_name, sep = "")
)

proj.env$our_db <- DBI::dbReadTable(
  proj.env$con,
  params$source_data_table_name
) %>%
  mutate(
    bordereaux_month = as.Date(bordereaux_month),
    v_created = as.Date(v_created),
    query_date = as.Date(query_date)
  )

DBI::dbDisconnect(proj.env$con)
```

Validate the data by checking for missing values in the key columns.

```{r, Check Overrides have matching keys}
# Check that all keys in the overrides table are present in the data
if (any(!proj.env$extract_dates$nre_feed %in% proj.env$our_db$nre_feed & !proj.env$extract_dates$treaty_year %in% proj.env$our_db$treaty_year)) {
  stop("Keys in the extract_dates table are not present in the data.")
}

if (any(!proj.env$experience_to_omit$nre_feed %in% proj.env$our_db$nre_feed & !proj.env$experience_to_omit$treaty_year %in% proj.env$our_db$treaty_year)) {
  stop("Keys in the experience_to_omit table are not present in the data.")
}
```

Apply Overrides to address data issues and filter data by Valuation Date

```{r, Apply Overrides and Filter Data by Valuation Date}
# Find the latest query_date.
# Store it as a string and filter to only retain matching rows
proj.env$latest_refresh_date <- as.Date(max(proj.env$our_db$v_created))

proj.env$our_db_filtered <- proj.env$our_db %>%
  dplyr::left_join(proj.env$extract_dates, by = c("nre_feed", "treaty_year")) %>%
  dplyr::left_join(proj.env$experience_to_omit, by = c("nre_feed", "treaty_year")) %>%
  dplyr::filter(
    v_created == coalesce(extract_to_use, proj.env$latest_refresh_date),
    bordereaux_month != coalesce(month_to_exclude, as.Date(format(Sys.Date() + lubridate::period(1, units = "months"), "%Y-%m-01")))
  )
```

Verify that all data is present and that the latest refresh date is correct.

```{r, Verify Data}
# Query the original data for rows matching the latest refresh date
proj.env$our_db_latest <- proj.env$our_db %>%
  filter(v_created == proj.env$latest_refresh_date)

# Find rows in the original data that are not in the filtered data
proj.env$missing_rows <- anti_join(
  proj.env$our_db_latest,
  proj.env$our_db_filtered,
  by = c("nre_feed", "treaty_year", "bordereaux_month")
)

# Check if the missing rows have a reason in the extract_dates or experience_to_omit tables
proj.env$missing_rows_with_reason <- proj.env$missing_rows %>%
  left_join(proj.env$extract_dates, by = c("nre_feed", "treaty_year")) %>%
  left_join(proj.env$experience_to_omit, by = c("nre_feed", "treaty_year")) %>%
  mutate(
    reason = coalesce(extract_to_use_reason, reason_to_drop)
  ) %>%
  select(
    nre_feed,
    treaty_year,
    bordereaux_month,
    v_created,
    reason
  )

# Print the missing rows with reasons
print(proj.env$missing_rows_with_reason)

if (
  nrow(proj.env$missing_rows)
  + nrow(proj.env$our_db_filtered)
  != nrow(proj.env$our_db_latest)
) {
  stop("Missing rows found in the data. Please investigate.")
}
```

Validate that any exceptions in the data have a matching key.

```{r, Validate Exceptions}
# Check that all keys in the exceptions table are present in the data
if (any(!proj.env$overrides_table$nre_feed %in% proj.env$our_db$nre_feed & !proj.env$overrides_table$treaty_year %in% proj.env$our_db$treaty_year & !proj.env$overrides_table$bordereaux_month %in% proj.env$our_db$bordereaux_month)) {
  stop("Keys in the exceptions table are not present in the data.")
}
```

Now that we have the latest data validated, we can correct any known exceptions in the data.

```{r, Correct Exceptions}
# Correct exceptions in the data
proj.env$our_db_filtered <- proj.env$our_db_filtered %>%
  left_join(proj.env$overrides_table, by = c("nre_feed", "treaty_year", "bordereaux_month")) %>%
  mutate(across(
    starts_with("tf_assumed_"),
    ~ coalesce(get(paste0("override_", sub("tf_assumed_", "", cur_column()))), .)
  )) %>%
  select(
    colnames(proj.env$our_db_filtered),
    reason_to_override
  )
```

Load the keys we use for projections

```{r, Translate Source Data to our Index Key}
# Read in translations from program names to program used by our organization
proj.env$key_translations <- read.csv(
  paste(params$reporting_path, params$key_map, sep = ""),
  stringsAsFactors = FALSE
)
```

Validate the key mappings by checking the source table.

```{r, Check Mappings have matching keys}
# Check that all keys to map exist in the treaty table
if (any(!proj.env$key_translations$nre_feed %in% proj.env$our_db$nre_feed & !proj.env$key_translations$treaty_year %in% proj.env$our_db$treaty_year)) {
  stop("Keys in the key_translations table are not present in the experience table.")
}
```

Consolidate (Aggregate) source data according to our keys

```{r, Join and Consolidate Data using Keys}
# Join the keys to the dataframe
proj.env$vel_df <- proj.env$our_db_filtered %>%
  left_join(proj.env$key_translations, by = c("nre_feed", "treaty_year"))
```

This step consolidates treaties that were broken out by our data provider into a single treaty.

```{r, Summarize Data by Keys}
# Rows are grouped by KEY and month.
# All numeric values are summed and NAs are preserved.
proj.env$vel_df1 <- proj.env$vel_df %>%
  group_by(KEY, bordereaux_month) %>%
  summarize(across(where(is.numeric), sum_with_na),
    across(where(~ is.numeric(.x)), first),
    .groups = "drop"
  )
```

Rename the columns and subtract salvage & subrogation from Paid Loss

```{r, Adjust the data for experience to date}
proj.env$vel_df2 <- proj.env$vel_df1 %>%
  rename(
    DATAWPARTMONTH = tf_assumed_gross_written_premium,
    DATAEPARTMONTH = tf_assumed_earned_premium,
    DATAPARTPAIDALAEMONTH = tf_assumed_paid_alae,
    DATAPARTCASEOSLOSS = tf_assumed_expected_losses
  ) %>%
  mutate(
    DATAPARTPAIDLOSSMONTH = tf_assumed_paid_losses - coalesce(tf_assumed_ss, 0),
    DATAPARTPAIDULAEMONTH = NA, # No data for paid ULAE (yet)
    DATAPARTCASEOSALAE = NA, # Outstanding ALAE Balance
    DATAPARTCASEOSULAE = NA, # Outstanding ULAE Balance
    MONTH = format(bordereaux_month, "%Y%m")
  )
```

Consolidate to only keeping relevant columns

```{r, Consolidate Data}
proj.env$vel_df3 <- proj.env$vel_df2 %>%
  select(
    KEY,
    MONTH,
    DATAWPARTMONTH, # written premium
    DATAEPARTMONTH, # earned premium
    DATAPARTPAIDLOSSMONTH,
    DATAPARTPAIDALAEMONTH,
    DATAPARTPAIDULAEMONTH,
    DATAPARTCASEOSLOSS,
    DATAPARTCASEOSALAE,
    DATAPARTCASEOSULAE
  )
```

Validate the keys in the experience table exist in the treaty position table.

```{r, Check experience keys match treaty position keys}
# Check that all keys to map exist in the treaty table
if (any(!proj.env$vel_df3$KEY %in% proj.env$treaty_positions$KEY)) {
  stop("Keys in the experience table are not present in the treaty position table.")
}
```

### Join and Prepare Data

This step creates a dataframe with every treaty having monthly rows from 2020 - 2070.

We also join the premium and loss data from DATATODATE, rename columns, and sort by treaty and month.

```{r, Prepare data}
# Create a vector with monthly dates from 2020 through 2070
proj.env$date_vector <- seq.Date(as.Date("2020-01-01"), as.Date("2070-12-01"), by = "month")

# expand.grid works like a cross_join, but doesn't require the vectors to be dataframes
proj.env$MonthlyProjections <- expand.grid(
  KEY = proj.env$treaty_positions$KEY,
  MONTH = proj.env$date_vector
) %>%
  left_join(proj.env$treaty_positions, by = c("KEY")) %>%
  left_join((proj.env$vel_df3 %>% mutate(MONTH = lubridate::ym(MONTH))), by = c("KEY", "MONTH")) %>%
  rename(
    Program_Name = Name,
    Carrier = Carrier,
    Primary_LOB = PrimaryLOB,
    Secondary_LOB = SecondaryLOB,
    Effective_Date = EffectiveDate,
    Expiration_Date = ExpirationDate,
    Treaty_Length = TreatyLength,
    Suffix = Suffix,
    RA_or_LOD = RAvLOD,
    Total_Subject_Premium = TotalSubjectPremium,
    Target_Participation = TargetParticipation,
    Inherited_UEPR = InheritedUEPROutsideParticipation,
    ALAE_ULAE_In_Ceded_LR = ULAEorALAE,
    Est_Payout_Duration_Yrs = EstimatedPayoutDuration,
    ULAE_Perc = ULAEOutsideCedingCommission,
    Reins_Broker_Perc = ReinsuranceBrokerCommission,
    Expenses_Perc = SeparateExpenses,
    Policy_Length = AssumedPolicyLengthMonths,
    WP_Monthly_Data = DATAWPARTMONTH,
    EP_Monthly_Data = DATAEPARTMONTH,
    Loss_Paid_Monthly_Data = DATAPARTPAIDLOSSMONTH,
    ALAE_Paid_Monthly_Data = DATAPARTPAIDALAEMONTH,
    ULAE_Paid_Monthly_Data = DATAPARTPAIDULAEMONTH,
    Loss_Case_Reserve_Data = DATAPARTCASEOSLOSS,
    ALAE_Case_Reserve_Data = DATAPARTCASEOSALAE,
    ULAE_Case_Reserve_Data = DATAPARTCASEOSULAE,
    # These should be removed and just read from Treaty table
    Prior_No_Impr_Gross_LALAE_Ratio_Excl_Corr = PriorNoImprovementGrossLALAERatio,
    Prior_Half_Impr_Gross_LALAE_Ratio_Excl_Corr = PriorHalfImprovementGrossLALAERatio,
    LALAE_Ratio_No_Improv = PriorNoImprovmentNetLALAERatio,
    LALAE_Ratio_Half_Improv = PriorHalfImprovmentNetLALAERatio,
    LALAE_Ratio_Break_Even = PriorBreakevenLALAERatio
  ) %>%
  # rowwise() %>%
  mutate(
    MaturityDateOfNewPolicies = as.Date(MONTH %m+% Policy_Length) - 1,
    MonthsNewPolicyInforceAfterTreatyEnds = lubridate::period(ifelse(
      MONTH >= Effective_Date & MONTH <= Expiration_Date,
      pmax(as.numeric(months_between(Expiration_Date, MaturityDateOfNewPolicies), "months"), 0),
      0
    ), units = "months"),
  ) %>%
  # ungroup() %>%
  arrange(KEY, MONTH)
```

### Project Monthly Earned Pattern against Inherited UEPR

This step projects a monthly earned pattern to the Inherited UEPR, resulting in the IUEPR_Monthly column.

This only applies to LOD treaties.

```{r, Project Monthly Earned Pattern against Inherited UEPR}
# Define a function used for projecting Inherited UEPR
# NL: For LOD policies only
#     LOD policies are comprised of two components,
#         1) New Risks, and
#         2) Inherited LOD from prior UW period
# Assumptions:
#     1) all inherited UEPR is on policies paid at policy inception
#     2) The average underlying policy is written mid month
# EFF: the effective date of the treaty (assumed to be the first of the month)
# PLEN: Policy Length in Months -- the assumed length of the underlying exposure
# K: the difference between the month in question and the effective date of the treaty
# When this function is used below, it uses column "Policy_Length" for that treaty's assumed policy length
Inherited_UEPR_Pattern <- function(EFF, CAL, PLEN) {
  k <- as.numeric(months_between(EFF, CAL), "months")
  # need to convert period back to integer form
  PLEN <- as.numeric(PLEN, "months")
  dplyr::if_else(k >= 0 & k <= (PLEN - 1), (2 / PLEN) * (1 - (k + 0.5) / PLEN), 0)
  # {(12-(k+0.5)*(12/PLEN))/(6*PLEN)}
  # k+0.5 adjusts for the mid-month bind
  # (k+0.5)/PLEN yields the proportion of the overall length that has already passed (Parallelogram horizontal distance)
  # 1-(k+0.5)/PLEN yields the portion of the overall length yet to pass (Parallelogram horizontal distance)
}



# Project IUEPR by month using Inherited_UEPR_Pattern function
proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>%
  mutate(
    IUEPR_Total = tidyr::replace_na(Inherited_UEPR, 0),
    IUEPR_Perc = Inherited_UEPR_Pattern(Effective_Date, MONTH, Policy_Length),
    IUEPR_Monthly = IUEPR_Total * IUEPR_Perc
  )
```

The following is a check to ensure that LOD Treaties have their sum of IUEPR_Monthly equal to IUEPR_Total.

```{r, Validate UEPR calc gets back to total}
proj.env$MonthlyProjections %>%
  distinct(KEY, .keep_all = TRUE) %>%
  summarize(Total_IUEPR = sum(IUEPR_Total)) %>%
  pull(Total_IUEPR)

sum(proj.env$MonthlyProjections$IUEPR_Monthly)
```

### Project Company's Share of Monthly Written Premium

This step projects share of monthly written premium and revises projections with reported data.

The projection evenly distributes the estimated total written premium across the n month treaty period.

```{r, Project Business Proportion of Written Premium}
# Define a function used for projecting monthly written premium
# Calculates the expected fraction of written premium in a given month by checking if that month
# is between the effective date and the expiration date
# EFF = The effective date of the policy, assumed to be the first of the month
# CAL = The calendar month for which we are calculating the expected written premium, assumed to be the first of the month
# TLEN = The length of the treaty in months
# Returns 1/TLEN if the month is within the treaty period, 0 otherwise
WrittenPattern <- function(EFF, CAL, TLEN) {
  treaty_month <- as.numeric(months_between(EFF, CAL), "months")
  # need to convert period back to integer form
  TLEN <- as.numeric(TLEN, "months")
  dplyr::if_else(treaty_month >= 0 & treaty_month < TLEN, 1 / TLEN, 0)
}

# Project Written Participation by Month
proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>%
  dplyr::mutate(
    # Our QS (quota share) of the Estimated Total Written Premium
    WP_Total = Total_Subject_Premium * TargetParticipationPercentage,
    # Allocate 1/n for all months during the treaty period (where n is the number of months the treaty is effective for), 0 otherwise
    WP_Perc = WrittenPattern(Effective_Date, MONTH, Treaty_Length),
    # Project monthly written participation
    WP_Monthly = WP_Total * WP_Perc,
  )

# Record the total difference between projected written premiums and those that were reported
# (this isolated step can be consolidated if there's no problem with adding an extra WP_Proj
# column to retain the original projected values)
proj.env$WP_Proj_vs_Rept <- proj.env$MonthlyProjections %>%
  dplyr::filter(!is.na(WP_Monthly_Data)) %>%
  mutate(WP_Proj_vs_Rept = WP_Monthly - WP_Monthly_Data) %>%
  summarize(WP_Proj_vs_Rept_Total = sum(WP_Proj_vs_Rept)) %>%
  pull(WP_Proj_vs_Rept_Total)

# Replace projected data with actual data where available
proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>%
  mutate(
    WP_Monthly = datareplace(WP_Monthly, WP_Monthly_Data),
    # Amount of Written Premium that will be earned in next treaty (LOD only)
    WP_ToEarnNextTreaty = ifelse(
      RA_or_LOD == "LOD",
      WP_Monthly * (as.numeric(MonthsNewPolicyInforceAfterTreatyEnds, "months") / as.numeric(Policy_Length, "months")),
      0
    )
  )
```

The following is a check to ensure the originally anticipated WP participation (WP_Total) carried forward through the calculations.

```{r, Validate Written Premium Participation}
# Sum of anticipated treaty WP participations
proj.env$MonthlyProjections %>%
  distinct(KEY, .keep_all = TRUE) %>%
  summarize(Total_WP_Participation = sum(WP_Total)) %>%
  pull(Total_WP_Participation)

# WP_Monthly contains projected monthly WP and, where available, reported WP
# Therefore, we sum the revised WP_Monthly column and add the monthly discrepancies between
# projected and actual reported values.
sum(proj.env$MonthlyProjections$WP_Monthly) + proj.env$WP_Proj_vs_Rept
```

### Project Company's Share of Monthly Earned Premium

This step projects share of monthly earned premium and revises projections with reported data. The calculation for monthly earned premiums uses the updated WP_Monthly from the previous section.

Two temporary tables are created to form and calculate the sumproduct of earned premium for each month, which are then joined back to MonthlyProjections.

```{r, Project Business Proportion of Earned Premium}
# Define a function used for used for projecting monthly earned premium
# k: month of treaty
# PLEN: Policy Length in Months - the assumed length of the underlying exposure (passed from "Policy_Length")
# The function generates a vector with proportions evenly distributed across the assumed months
# of policy length. The 1st month, however, splits half of its share with the month following the policy duration.
EarnedPattern <- function(k, PLEN) {
  # need to convert period back to integer form
  k <- as.numeric(k, "months")
  PLEN <- as.numeric(PLEN, "months")
  case_when(
    (k >= 1 & k <= (PLEN - 1)) ~ 1 / PLEN,
    (k == 0 | k == PLEN) ~ 1 / (2 * PLEN),
    TRUE ~ 0
  )
}

# Create first table containing all rows in MonthlyProjections dataframe
proj.env$table1_Earning_Pattern <- proj.env$MonthlyProjections %>%
  select(
    # Keys
    Program_Name,
    Primary_LOB,
    Effective_Date,
    Suffix,
    WP_Month = MONTH,
    # Attribute
    RA_or_LOD,
    Policy_Length,
    WP_Monthly,
    Expiration_Date,
  )

# Create a second table by joining on the newly created EP_Month, resulting in multiple earning months for each WP month
proj.env$table2_Earning_Pattern <- left_join(
  proj.env$table1_Earning_Pattern,
  select(proj.env$table1_Earning_Pattern, EP_Month = WP_Month) %>% distinct(),
  join_by(WP_Month <= EP_Month)
) %>%
  dplyr::mutate(
    # Calculate the lag between writing and earning
    Earned_Lag = months_between(WP_Month, EP_Month),
    # Calculate the percent of premium earned in each month from each WP month
    Earned_Pct = ifelse(
      RA_or_LOD == "LOD" & EP_Month > Expiration_Date,
      0,
      EarnedPattern(Earned_Lag, Policy_Length)
    ),
    # Earned Premium from each WP month is the product of that month's WP
    EP_Monthly = Earned_Pct * WP_Monthly,
  ) %>%
  # Remove all rows that don't have an Earned Premium. (And retain negative earned premiums.)
  dplyr::filter(EP_Monthly != 0) %>%
  # Sum across WP months because it's irrelevant which WP month it was sourced from, rather only when it was earned
  group_by(
    Program_Name,
    Primary_LOB,
    Effective_Date,
    Suffix,
    EP_Month
  ) %>%
  summarize(EP_Monthly = sum(EP_Monthly))
# ungroup() # ungroup shouldn't be needed because summarize will have collapsed the rows by the group_by columns

# Join the earned premiums (EP_Monthly) to MonthlyProjections
proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>%
  left_join(select(proj.env$table2_Earning_Pattern, c(Program_Name, Primary_LOB, Effective_Date, Suffix, EP_Month, EP_Monthly)),
    by = c(
      "Program_Name",
      "Primary_LOB",
      "Effective_Date",
      "Suffix",
      "MONTH" = "EP_Month"
    )
  ) %>%
  mutate(EP_Monthly = tidyr::replace_na(EP_Monthly, 0))


proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>%
  mutate(
    # Projected Earned Participation by Month = Inherited UEPR + Exposure Policy Earned Premium
    EP_Monthly_Incl_IUEPR = IUEPR_Monthly + EP_Monthly,
    # Swapping out Projected EP for Actual EP where available
    # [NOTE] doing this means that WP_To_Date will not be equal to EP_To_Date in the tail months
    # which means the UEPR calc below will not converge to 0 in all cases
    # TODO: this is a known issue and we need to develop a fix for it
    EP_Monthly_Incl_IUEPR = datareplace(EP_Monthly_Incl_IUEPR, EP_Monthly_Data),
  ) %>%
  group_by(KEY) %>%
  mutate(
    # Total Earned Premium To Date (for each treaty)
    EP_To_Date = cumsum(EP_Monthly_Incl_IUEPR),
    # Total Written Premium To Date (for each treaty)
    WP_To_Date = cumsum(WP_Monthly),
    # Calculate the UEPR; note this will not be zero at end for LOD treaties
    UEPR = IUEPR_Total + WP_To_Date - EP_To_Date,
    # Actual UEPR Balance remaining on IUEPR alone
    # IUEPR_Balance = IUEPR_Total - cumsum(IUEPR_Monthly),
    # Total Written Premium to Disinherit as it will runoff to the next treaty,
    DUEPR_Balance = cumsum(WP_ToEarnNextTreaty)
  ) %>%
  ungroup()
```

```{r, Validate Written to Earned Premium}
# Check that total earned premium == total written premium
sum(proj.env$MonthlyProjections$WP_Monthly)
sum(proj.env$MonthlyProjections$EP_Monthly)
```

### Undeveloping Earned Premiums for Projecting Paid and Reported Losses

This step projects paid and reported losses by undeveloping the total earned premium (including Inherited UEPR where relevant).

Two temporary tables are used to form and calculate the sumproducts of the losses that occur for each month, which are then joined back to MonthlyProjections.

```{r, Validate Keys in Dev Factors}
# Check that all keys in the development factors table are present in the data
if (any(!proj.env$development_factors$KEY %in% proj.env$MonthlyProjections$KEY)) {
  proj.env$missing_keys <- unique(proj.env$development_factors$KEY[!proj.env$development_factors$KEY %in% proj.env$MonthlyProjections$KEY])
  warning(paste(
    "The following keys in the development factors table are not present in the projection table:\n",
    paste(proj.env$missing_keys, collapse = "\n")
    )
  )
}

# Check that all keys in the projection table are present in the development factors
if (any(!proj.env$MonthlyProjections$KEY %in% proj.env$development_factors$KEY)) {
  proj.env$missing_keys <- unique(proj.env$MonthlyProjections$KEY[!proj.env$MonthlyProjections$KEY %in% proj.env$development_factors$KEY])
  warning(paste(
    "The following keys in the projection table are not present in the development factors table:\n",
    paste(proj.env$missing_keys, collapse = "\n")
    )
  )
}
```

```{r, Apply development factors}
# Prepare the development factors dataset to have 0 months lag indicating the current month
proj.env$DEVS_rev <- proj.env$development_factors %>%
  arrange(KEY, LAG_ACCMONTH) %>%
  dplyr::filter(LAG_ACCMONTH != 0) %>% # Remove rows where LAG_ACCMONTH is 0
  mutate(LAG_ACCMONTH = lubridate::period(LAG_ACCMONTH - 1, unit = "months")) # Decrement LAG_ACCMONTH by 1

# Create an initial table that will be used for calculating losses
proj.env$t_Loss <- proj.env$MonthlyProjections %>%
  select(
    # Primary Key
    KEY,
    # Unique Key
    Program_Name,
    Primary_LOB,
    Effective_Date,
    Suffix,
    Effective_Date,
    MONTH,
    # Attribute
    EP_Monthly_Incl_IUEPR
  ) %>%
  # This is basically the same as using window functions in SQL
  # e.g. SUM(x) OVER (PARTITION BY y) in this case
  group_by(KEY) %>%
  mutate(Total_Earned_Partic = sum(EP_Monthly_Incl_IUEPR)) %>%
  ungroup()

# Join the development factors
proj.env$t_Loss_2 <- proj.env$t_Loss %>%
  full_join(proj.env$DEVS_rev, by = c("KEY"), relationship = "many-to-many") %>%
  dplyr::mutate(
    # This column contains the month for the paid and reported losses
    LAG_date = MONTH %m+% LAG_ACCMONTH,
    # Paid and Reported losses using development factors on Earned Premium
    # I renamed Paid_Loss -> EP_Undev_Paid
    # I renamed Reported_Loss -> EP_Undev_Reported
    EP_Undev_Paid = EP_Monthly_Incl_IUEPR * PAID_PERCENT,
    EP_Undev_Reported = EP_Monthly_Incl_IUEPR * REPORTED_PERCENT
  ) %>%
  # Remove the rows that are past our data's time frame
  dplyr::filter(LAG_date <= as.Date("2070-12-01")) %>%
  # Sum across loss months because it's irrelevant which EP month it was sourced from,
  # rather only when the loss is paid/reported
  group_by(
    Program_Name,
    Primary_LOB,
    Effective_Date,
    Suffix,
    Total_Earned_Partic,
    LAG_date
  ) %>%
  summarize(
    EP_Undev_Paid = sum(EP_Undev_Paid),
    EP_Undev_Reported = sum(EP_Undev_Reported)
  ) %>%
  # ungroup() %>% # ungroup shouldn't be needed because summarize will have collapsed the rows by the group_by columns
  # either that or a new group by treaty should be used?
  # Ensure all months past the scope of the development factors have the full paid and reported amounts
  mutate(
    EP_Undev_Paid = ifelse(
      as.numeric(months_between(Effective_Date, LAG_date), "months") >= 240,
      Total_Earned_Partic,
      EP_Undev_Paid
    ),
    EP_Undev_Reported = ifelse(
      as.numeric(months_between(Effective_Date, LAG_date), "months") >= 240,
      Total_Earned_Partic,
      EP_Undev_Reported
    )
  ) %>%
  ungroup() %>% # ungroup required here because it won't allow removing a column in the group_by list above
  # Remove columns that won't be imported
  select(-Total_Earned_Partic)

proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>% left_join(proj.env$t_Loss_2, by = c(
  "Program_Name",
  "Primary_LOB",
  "Effective_Date",
  "Suffix",
  "MONTH" = "LAG_date"
))
```

The following is a check to ensure the columns for paid and reported losses sum to the total undeveloped earned premium. There may be slight deviances due to the different calculations involved with each field.

```{r, Validate Losses}
proj.env$Check_EP_Undev_Losses <- proj.env$MonthlyProjections %>%
  group_by(KEY) %>%
  summarize(
    EP_Monthly_Incl_IUEPR_Total = sum(EP_Monthly_Incl_IUEPR),
    EP_Undev_Paid_Total = max(EP_Undev_Paid),
    EP_Undev_Reported_Total = max(EP_Undev_Reported)
  )

sum(proj.env$Check_EP_Undev_Losses$EP_Monthly_Incl_IUEPR_Total)
sum(proj.env$Check_EP_Undev_Losses$EP_Undev_Paid_Total)
sum(proj.env$Check_EP_Undev_Losses$EP_Undev_Reported_Total)
```

### Create Additional Projections

This step calculates the dollar amounts for ULAE, reinsurance broker commission, expenses, and multiple LALAE values.

```{r, Additional Projections}
proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>%
  mutate(
    ULAE_Total = EP_To_Date * ULAE_Perc,
    Reins_Broker_Comm_Total = EP_To_Date * Reins_Broker_Perc,
    Expenses_Total = EP_To_Date * Expenses_Perc,
    # These should be removed and just read from Treaty table
    LALAE_Paid_Total_No_Improv = EP_Undev_Paid * LALAE_Ratio_No_Improv,
    LALAE_Paid_Total_Half_Improv = EP_Undev_Paid * LALAE_Ratio_Half_Improv,
    LALAE_Paid_Total_Break_Even = EP_Undev_Paid * LALAE_Ratio_Break_Even,
    LALAE_Reported_Total_No_Improv = EP_Undev_Reported * LALAE_Ratio_No_Improv,
    LALAE_Reported_Total_Half_Improv = EP_Undev_Reported * LALAE_Ratio_Half_Improv,
    LALAE_Reported_Total_Break_Even = EP_Undev_Reported * LALAE_Ratio_Break_Even,
    LALAE_Reserve_No_Improv = LALAE_Ratio_No_Improv * (EP_To_Date - EP_Undev_Paid),
    LALAE_Reserve_Half_Improv = LALAE_Ratio_Half_Improv * (EP_To_Date - EP_Undev_Paid),
    LALAE_Reserve_Break_Even = LALAE_Ratio_Break_Even * (EP_To_Date - EP_Undev_Paid),
    LALAE_IBNR_No_Improv = LALAE_Ratio_No_Improv * (EP_To_Date - EP_Undev_Reported),
    LALAE_IBNR_Half_Improv = LALAE_Ratio_Half_Improv * (EP_To_Date - EP_Undev_Reported),
    LALAE_IBNR_Break_Even = LALAE_Ratio_Break_Even * (EP_To_Date - EP_Undev_Reported)
  )
```

### Export Monthly Projections csv

```{r, Export Monthly Projections}
# Export the Monthly Projections to a csv file

# Get today's date in YYYYMMDD format
proj.env$today <- format(Sys.Date(), "%Y%m%d")

# Create the file name
proj.env$file_name <- paste0(
  params$reporting_path,
  proj.env$today, "_",
  format(as.Date(params$date_of_valuation, "%Y-%m-%d"), "%Y%m%d"),
  "_VAL_MONTHLY_PROJECTIONS.csv"
)

# Export the Monthly Projections to a csv file with the dynamic file name
write.csv(
  proj.env$MonthlyProjections,
  file = proj.env$file_name,
  row.names = FALSE
)
```
